### 2021-04-06 chapters/introduction.tex
Initial Overleaf Import

\chapter{Introduction}
\documentclass[../main]{subfiles}
\begin{document}

\label{ch:introduction}
Context: what is the bigger scope of the problem you are trying to solve? Try to connect to societal/economical challenges.
Problem Analysis: Here you present your analysis of the problem situation that your research will address.
How does this problem manifest itself at your host organisation?
Also summarises existing scientific insight into the problem.
\glo

\section{Problem statement}

\subsection{Research questions}
To tackle these issues, we investigate ...

\subsection{Research method}

\section{Contributions}
Our research makes the following contributions:
\begin{enumerate}
	\item 1
	\item 2
	\item 3
\end{enumerate}

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. 
Chapter~\ref{ch:research} describes ... 
Results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work}, contains the work related to this thesis.
Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

\end{document}