### 2021-04-13 acronym.tex
Added introduction + first draft of rq1.1

%helpful acronym examples
\newacronym{poc}{POC}{proof-of-concept}

\newacronym{sut}{SUT}{System Under Test}

\newacronym{strew}{STREW}{Software Testing and Reliability Early Warning}

\newacronym{sat}{SAT}{Software Analysis Toolkit}

\newacronym{tloc}{TLOC}{Lines Of Test Code}

\newacronym{loc}{LOC}{Lines of Code}

\newacronym{taime}{TAIME}{Test Suite Assessment and Improvement Method}

\newacronym{tqm}{TQM}{Test Quality Model}

\newacronym{tdd}{TDD}{Test-Driven Development}

\newacronym{cor}{COR}{Conditional Operator Replacement}

\newacronym{ror}{ROR}{Relational Operator Replacement}

\newacronym{ast}{AST}{Abstract Syntax Tree}