### 2021-04-13 chapters/4_mutation_tool.tex

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Mutation testing tools}
\label{ch:mutation-tool}

Before we can start clustering mutants we need to select a tool that helps us generate mutants. 
Our goal is to find the tool that generates the greatest amount of mutants. 
\newline
First, we review existing literature. 
Secondly we extend the comparison by adding information about the current state of the tools.
With the combination of the information gathered we make a decision on what tool to use for this thesis.

\section{State of the art documented}
\label{ch:tooling-literature}
The literature used several criteria for comparing mutation testing tools for Java:
\begin{itemize}
  \item Execution time of each tool.
  \item Tool's application cost in terms of the number of test cases needs to be generated and the the number of equivalent mutants that would have to be inspected.
  \item Effectiveness of the mutation adequate test suite of each tool. Each test of the test suite contributes to the
effectiveness score, i.e., if one test is removed, less than 100\% effectiveness score is achieved. A
cross-testing technique is applied to evaluate the effectiveness each tool’s mutation adequate test
suite. The adequate test suite of each tool is run on the set of mutants generated by each other tool.
\end{itemize}
Kintis et al. analysed and compared the effectiveness of PIT,  MuJava and Major\cite{Kintis2016AnalysingStudy}. 
They started out with performing a mini literature survey on mutation testing tools used in research on test effectiveness in Java. 
With a data set of twelve methods and six Java projects they evaluated each tool using a cross-testing technique.
They found that the mutation adequate test suite of  MuJava was the most effective, followed by Major and PIT.
The score of the application cost was the inverse of the previous ranking: PIT required the least test cases and generated the smallest set of equivalent mutants.
During their research they also compared the different mutant operators the tools supported. 
An overview of mutant operators per tool was created. 
They compared the mutant operators and reported overlapping operators.
\newline
Marki and Lindstrom performed the research on the same set of mutation tools\cite{Marki2017MutationJava}. 
Similar to Kintis et al., they applied the same cross-testing technique for three small Java programs. 
These programs are often used in testing literature.
They found that none of the mutation tools subsume \todo{add subsumtion to word list} each other. 
If the tools were ranked,  MuJava would generate the strongest mutants followed by Major and PIT.
Furthermore they found that  MuJava generated significantly more equivalent mutants.
MuJava also had an execution time twice the amount of that of Major and PIT combined.
\newline
Laurent et al., introduced PIT+, an version of PIT with extended set of mutantors\cite{Laurent2017AssessingPIT}.
They used the tests suites generated by kintis et al., and combined these into an adequate test suite that would detect the combined set of mutants generated by PIT, muJava and Major. 
They found that the set of mutants generated by PIT+ was equally strong as the combined sets of mutants of all three tools. 
The extended set of mutants used in PIT+ is now integrated in the PIT project\cite{pit+}.

\section{State of the art tools}
The literature is clear about how the mutation testing tools compare to each other.
However the most recent study is from 2017.
At the time of writing this thesis the studies are at least four years old. 
Some of the tools have been developed further or are still actively developed\cite{pit-releases,Major}.
\newline
There are three candidate tools; Major,  MuJava and PIT.
We extended the existing mutant operator comparison for these tools with the operators that have been added since the publishing date of the literature.
We also researched the overlap for the new mutant operators.
\newpage
Table \ref{tab:mutators-current} shows the mutant operators that were developed after the publication of the literature. The amount of mutant operators for  MuJava and PIT have increased significantly while there were no new mutant operators for Major.
\begin{table}[htb]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{ MuJava}}                           & \multicolumn{2}{|c|}{\textbf{PIT}}                      \\ \hline
IHD & Hiding variable deletion                                  & EM                    & Empty Returns                   \\ \hline
IHI & Hiding variable   insertion                               & FR                    & False Returns                   \\ \hline
IOD & Overriding method   deletion                              & NR                    & Null Returns                    \\ \hline
IOP & Overriding method   calling position change               & TR                    & True Returns                    \\ \hline
IOR & Overriding method   rename                                & PR                    & Primitive Returns               \\ \hline
ISI & super keyword insertion                                   & ER                    & Experimental Switch             \\ \hline
ISD & super keyword deletion                                    & BI                    & Big Integer                     \\ \hline
IPC & Explicit call to a   parent’s constructor deletion        & NRC                   & Naked Receiver                  \\ \hline
PNC & new method call with   child class type                   & N                     & Negation                        \\ \hline
PMD & Member variable   declaration with parent class type      & AOR                   & Arithmic Operator Replacement   \\ \hline
PPD & Parameter variable   declaration with child class type    & AOD                   & Arithmic Operator Deletion      \\ \hline
PCI & Type cast operator   insertion                            & CR                    & Constant replacement            \\ \hline
PCC & Cast type change                                          & BO                    & Bitwise Operator                \\ \hline
PCD & Type cast operator   deletion                             & ROR                   & Relational Operator Replacement \\ \hline
PRV & Reference assignment   with other comparable variable     & UOI                   & Unary Operation Insertion       \\ \hline
OMR & Overloading method   contents replace                     & \multicolumn{1}{l|}{} &                                 \\ \hline
OMD & Overloading method   deletion                             & \multicolumn{1}{l|}{} &                                 \\ \hline
OAC & Arguments of   overloading method call change             & \multicolumn{1}{l|}{} &                                 \\ \hline
JTI & this keyword insertion                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
JTD & this keyword deletion                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
JSI & static modifier   insertion                               & \multicolumn{1}{l|}{} &                                 \\ \hline
JSD & static modifier   deletion                                & \multicolumn{1}{l|}{} &                                 \\ \hline
JID & Member variable   initialization deletion                 & \multicolumn{1}{l|}{} &                                 \\ \hline
JDC & Java-supported default   constructor deletion             & \multicolumn{1}{l|}{} &                                 \\ \hline
EOA & Reference assignment   and content assignment replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
EOC & Reference comparison   and content comparison replacement & \multicolumn{1}{l|}{} &                                 \\ \hline
EAM & Acessor method change                                     & \multicolumn{1}{l|}{} &                                 \\ \hline
EMM & Modifier method change                                    & \multicolumn{1}{l|}{} &                                 \\ \hline
\end{tabular}
\caption{\label{tab:mutators-current}Current mutators supported.}
\end{table}
\newpage
Tools sometimes implement the same mutator differently, resulting in different mutant sets. Table \ref{tab:mutators-overlap} shows an overview over the relations between the mutators of MuJava and PIT. This overview takes into account the mutators that are not already in the studies of Kintis et al., and Marki and Lindstrom. However some of the new mutators map to the existing mutators.

\begin{table}[htb]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{ MuJava mutator}}      & \multicolumn{2}{c|}{\textbf{PIT mutator}}                         \\ \hline
AORB & Arithmetic Operator Replacement Binary      & \multirow{4}{*}{M}   & \multirow{4}{*}{Math}                      \\ \cline{1-2}
ASRS & Short-Cut Assignment   Operator Replacement &                      &                                            \\ \cline{1-2}
SOR  & Shift Operator   Replacement                &                      &                                            \\ \cline{1-2}
COR  & Conditional Operator   Replacement          &                      &                                            \\ \hline
AOIU & Arithmetic Operator   Insertion Unary       & \multirow{2}{*}{UOI} & \multirow{2}{*}{Unary Operation Insertion} \\ \cline{1-2}
AOIS & Arithmetic Operator   Insertion Short-cut   &                      &                                            \\ \hline
ODL  & Operator Deletion                           & AOD                  & Arithmic Operator   Deletion               \\ \hline
AODS & Arithmetic Operator   Deletion Short-cut    & RI                   & Remove Increments                          \\ \hline
ROR  & Relational Operator   Replacement           & CB                   & Conditionals Boundary                      \\ \hline
AORS & Arithmetic Operator   Replacement Short-Cut & I                    & Increments                                 \\ \hline
COD  & Conditional Operator   Deletion             & RC                   & Remove Conditionials                       \\ \hline
\end{tabular}
\caption{\label{tab:mutators-overlap}Overview of mutator overlaps between  MuJavav and PIT}
\end{table}
Next we tried to generate mutants with all three tools. 
All of the sources that we used for generating mutants are build with the build tools Maven or Gradle. 
Marki and Lindstrom\cite{Marki2017MutationJava} mention that Major supports Maven, but it is not documented anywhere.
For this reason we could not generate mutants with Major.
Table \ref{tab:mutants-generated} shows the amount of mutants generated per project and per tool. The results show that in all of the cases PIT generates more mutants than MuJava. In some cases the difference is significant. This difference can partially be explained by the fact that MuJava does not support source projects with Java version 1.7 or higher\cite{mujava}. All the the source code that uses features from this version of Java and higher resulted in an error during mutant generation.

\begin{table}[htb]
\centering
\begin{tabular}{|l|l|c|c|c|}
\hline
                                & \textbf{version} & \textbf{\begin{tabular}[c]{@{}c@{}}LOC\\ (Lines of code)\end{tabular}} & \multicolumn{1}{l|}{\textbf{MuJava}} & \multicolumn{1}{l|}{\textbf{PIT}} \\ \hline
org.apache.commons.numbers.core & 1.0-beta1        & 450                                                                    & 988                                  & 4384                              \\ \hline
Jodatime                        & 2.10.10          & 28811                                                                  & 52925                                & 112772                            \\ \hline
Zxing 3.4.1                     & 3.4.1            & 24792                                                                  & 65983                                & 161409                            \\ \hline
Google Auto Common              & 0.11             & 2338                                                                   & 20                                   & 5219                              \\ \hline
Google Auto Factory             & 1.0-beta9        & 1507                                                                   & 69                                   & 5832                              \\ \hline
Google Auto Value               & 1.7.5            & 7466                                                                   & 745                                  & 16746                             \\ \hline
Google Auto Service             & 1.0-rc6          & 750                                                                    & 0                                    & 475                               \\ \hline
ScribeJava-Core                 & 8.1.0            & 5709                                                                   & 1358                                 & 5746                              \\ \hline
Checkstyle                      & 8.41.1           & 38491                                                                  & 4529                                 & 100952                            \\ \hline
fastjson                        & 1.2.75           & 43405                                                                  & 70577                                & 116188                            \\ \hline
\end{tabular}
\caption{\label{tab:mutants-generated}Amount of mutants generated by MuJava and PIT}
\end{table}
\newpage
\section{Mutation tool}
We have studied literature on three different tools and extended the research with more recent data.
MuJava is not actively maintained and has not been updated in the last few years.
It does not support JUnit 4 and all versions of TestNG\cite{mujava}.
These are test engines\todo{add to vocabulary} and are crucial for executing tests.
MuJava also does not support source projects with java version 1.7 or higher\cite{mujava}.
Conforming to these requirements would decrease the set of projects we could use in or experiments.
\newline
While Major does support JUnit 4,
we could not generate mutants with this tool.
It would be too time consuming to customize all source projects to work with Major.
Furthermore there is not much documentation available for this tool.
\newline
PIT targets industry, is open source and actively maintained\cite{Kintis2016AnalysingStudy}.
It supports a wide scale of build tooling and is significantly faster than the other tools.
For example PIT provides a plugin system in which you can inject your own code in various stages of mutation testing\cite{pit}.
PIT also generated significantly more mutants in every project we have tested.
Based on these observations we decided that PIT would be the best choice for generating and executing mutants.
\end{document}