### 2021-04-19 acronym.tex
Added writeup for clustering algorithms + created chapter for rq1

%helpful acronym examples
\newacronym{poc}{POC}{proof-of-concept}

\newacronym{sut}{SUT}{System Under Test}

\newacronym{tloc}{TLOC}{Lines Of Test Code}

\newacronym{loc}{LOC}{Lines of Code}

\newacronym{cli}{CLI}{Command Line Interface}