### 2021-04-19 chapters/5_clustering_similarity.tex

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Clustering on similarity}\todo{do we add dissimilarity?}
\label{ch:clustering_similarity}
The similarity between a mutant and the original code it is mutated from can be defined as a characteristic. 
We can then cluster on similarity. 
In this chapter we implement a hierarchical clustering algorithm to cluster on the similarity of a mutant.

\section{Similarity measure}

\section{Hierarchical clustering}
\end{document}