### 2021-05-28 chapters/6_experiment_design.tex

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Experiment design}
\todo{expand this}
We hypothesise that there is a we can cluster on the characteristics of a mutant such that each mutant should be a representative for all of the mutants in its cluster.
Our goal is to achieve a mutation score that is as close as possible to the mutation score of a full set of executed mutants.

We design an experiment in which we implement the following steps:
\begin{enumerate}
    \item Extract all the characteristics.
    \item For each mutant calculate the Levenshtein distance.
    \item Cluster the mutants based on the characteristics.
    \item Randomly select one mutant from each cluster.
    \item Execute the selected mutants in PIT and calculate the mutation score.
\end{enumerate}
We elaborate on these steps in the following subsections.

\subsection{Levenshtein distance on Java byte code}
A mutant is a piece of code that differs in a predefined way from its parent.
Java source code compiles into Java byte code.
During the compilation of Java code certain context is abstracted away\cite{byteCodeEngineering}.
Optimisations are applied which also changes the Java byte code\cite{byteCodeEngineering}.
As a result the textual similarity between a Java code mutant and its parent is different from the similarity between a Java byte code mutant and its parent.
While the textual similarity may be different between the Java code and Java byte code the functionality remains the same.
Java byte code reflects more of the semantic nature of the source code than the source code itself does. 
\newline
PIT generates and executes mutants on byte code level\cite{pitestBytecode}. 
In other words PIT executes the unit tests of a source against the byte code of a mutant.
\newline
For this experiment we will use the Levenshtein distance between a Java byte code mutant and its parent.
By calculating the distance on byte code we filter out the context that is present in Java code.
This gives us a Levenshtein distance that represents more of the semantic difference, between a mutant and parent, than calculating the distance for Java code.

\subsection{Levenshtein distance implementation}
write here that we use a memory efficient algorithm

\subsection{Hierarchical clustering}
Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011HierarchicalStudy, Gothai2010PerformanceAlgorithms}. 
Hierarchical clustering is subdivided into agglomerative and divisive. 
The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
Hierarchical clustering uses different metrics which measures the the distance between two clusters and the linkage criteria\cite{Rajalingam2011HierarchicalStudy}. 
The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011HierarchicalStudy}.
\newline
Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
\newline
We identify each Levenshtein distance between a mutant and its parent as a separate cluster.
Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
The agglormorative hierarchical clustering algorithm with ward linkage will cluster our mutants represented as Levenshtein distances.
\todo{write about implementation}

\subsection{Weighed mutation score}
Our hypothesis states that each mutants should be a representative for all the mutants in its cluster. 
This translates to if a mutant survives or is killed all mutants in that cluster should survive or be killed respectively.
To reflect this in our experiment a weighed mutation score is calculated for the clustered mutants.
This weighed mutation score is the product of the result of a mutant(1 for killed and 0 for survived) and the amount of mutants in the cluster it represents.
The weighed mutation score is then comparable to the score of a full set as the total number of mutants will be the same.
\newline
For example, take a full set with a score of 75/100 killed mutants. \todo{example too much?}
This gives us a mutation score of 75\%. 
We then cluster the mutants in four clusters consisting of 12, 30, 38 and 20 mutants, respectively.
We randomly select four mutants of each cluster and execute them.
The mutants representing cluster one and four survive.
If we calculate the weighed score we get 68/100 which is 68\%.
We can then compare this to the score of a full set because the amount of mutants executed is the same: 75/100(75\%) and 68/100(68\%).

\subsection{PIT plugin system}
PIT offers a plugin system in which developers can inject their own functionality in PIT\cite{pitestPlugin}. 
There are two main types of plugins; a Mutation Result Listener and a Mutation interceptor\cite{pitestPlugin}.
A mutation result listener  receives the details of analysed mutations as they arrive\cite{pitestPlugin}.
A Mutation interceptor passed a complete list of all mutation that will be generated to each class before the mutation are challenged by tests\cite{pitestPlugin}.
The implementation of our experiment makes use of this plugins system.
\newline
We developed three separate plugins for our experiment.
The first plugin is a Mutation interceptor.
It analyses each mutant, calculates the Levenshtein distance and writes a mutant identifier and distance to a \acrshort{csv} file. 
It also clears the list of mutants provided by PIT to prevent execution. 
The second plugin is also a Mutation interceptor.
It reads the clustered mutants from a \acrshort{csv} file and compares them to the list of mutants provided by PIT.
Only the clustered mutants are passed back to PIT for execution.
The last plugin is a mutation result listener. 
This plugin calculates the weighed mutation score and outputs it to a text file.
\newline
The plugins in PIT are executed based on type and per source class.
This means that a plugin will never receive all the mutants generated for a whole project during execution.
Because of how PIT executes its plugins we have to execute PIT twice.
In the first run we calculate the Levenshtein distance for all mutants and in the second run we execute the selected mutants and calculate the weighed mutation score.
\end{document}