### 2021-06-23 chapters/5_research_question_1.tex
redo structure

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Research question 1}
\label{ch:reserach_question1}
\subfile{rq1/1_identifying_characteristics}
\subfile{rq1/2_experiment_design}
\subfile{rq1/3_results}
\end{document}