### 2021-06-23 chapters/6_research_question_2.tex

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Research question 2}
\label{ch:reserach_question2}
\subfile{rq2/1_experiment_design}
\end{document}