### 2021-07-03 chapters/7_discussion.tex

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Discussion}
\label{ch:discussion}
In this chapter, we discuss the results of our experiment(s) on ...

\begin{finding}
	Highlight like this an important finding of your analysis of the results.
	\label{find:important1}
\end{finding}

Refer to Finding~\ref{find:important1}.

\section{Threats to validity}
What could affect the validity of your research? Think of pitfalls of your research method, experimental setup, interpreting the results.

\end{document}