### 2021-07-03 chapters/8_related_work.tex

\documentclass[../main]{subfiles}
\begin{document}

\chapter{Related work}
\label{ch:related_work}
We divide the related work into the following categories:  ...

\end{document}