### 2021-07-14 acronym.tex
changes

%helpful acronym examples
\newacronym{poc}{POC}{proof-of-concept}

\newacronym{sut}{SUT}{System Under Test}

\newacronym{tloc}{TLOC}{Lines Of Test Code}

\newacronym{loc}{LOC}{Lines of Code}

\newacronym{cli}{CLI}{Command Line Interface}

\newacronym{csv}{CSV}{comma-separated values}

\newacronym{som}{SOM}{self organizing map}

\newacronym{fcm}{FCM}{fuzzy c-means}