### 2021-07-17 appendix/appendix.tex
WIP structure

\begin{appendices}

\chapter{PIT maven code}
\label{ap:PIT_maven_code}
Example of code used to execute a full set of mutants with PIT.
\begin{lstlisting}
    <plugins>
      <plugin>
        <groupId>org.pitest</groupId>
        <artifactId>pitest-maven</artifactId>
        <version>1.6.4</version>
        <dependencies>
          <dependency>
            <groupId>com.niverhawk</groupId>
            <artifactId>pitest-clustering-plugin</artifactId>
            <version>1.0-SNAPSHOT</version>
          </dependency>
        </dependencies>
        <configuration>
          <exportLineCoverage>true</exportLineCoverage>
          <mutators>
            <mutator>ALL</mutator>
          </mutators>
          <threads>6</threads>
        </configuration>
      </plugin>
\end{lstlisting}

\end{appendices}