### 2021-07-17 chapters/6_research_question_2.tex

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Black box appraoch}
\label{ch:reserach_question2}
\subfile{rq2/1_experiment_design}
\subfile{rq2/2_results}
\end{document}