### 2021-07-17 chapters/5_research_question_1.tex

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Research question 1}
\label{ch:reserach_question1}
The following sections describe the white box approach for clustering mutants.
We perform a qualitative analysis on the mutants to identify characteristics. 
These characteristics represent the mutants.
With the mutants represented we devise a methodology for clustering these mutants.
Our goal is to cluster every mutant that is generated while main tainting effectiveness.
\subfile{rq1/1_identifying_characteristics}
\subfile{rq1/2_clustering_mutants}
\end{document}