### 2021-07-17 chapters/8_results.tex

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Results}
\label{ch:results}
In this chapter we present the results of our experiments.
We divide the results in two sections. 
The first section displays the results of clustering with hierarchical clustering.
The second section displays the results of clustering with the \acrshort{fcm} model.
\newline
During the process of extracting the characteristics we found that calculating the Levenshtein distance took a long time.
Calculating the Levenshtein distance for the four biggest projects in the data set cost days.
Even for the smaller projects the extraction of the characteristics cost hours.
Including the Levenshtein distance defeats the purpose of increasing performance as every bit of performance we increase would be replaced by the cost of calculating the Levenshtein distance.
To see if we could mitigate this risk we decided to repeat the experiment without the Levensthein distance characteristic.

\subfile{results/1_results_rq1}
\subfile{results/2_results_rq2}
\end{document}