### 2021-07-20 chapters/6_research_question_2.tex
pong

\documentclass[../main]{subfiles}
\begin{document}
\chapter{Quantitative clustering approach}
\label{ch:reserach_question2}
In the following sections we devise a black box approach for clustering mutants.
We make use of the characteristics identified in Chapter \ref{ch:identifying_characteristics}.
While our goal remains the same as in Chapter \ref{ch:reserach_question1}, the way to achieve it is different.
We start by selecting a machine learning algorithm.
Next we select a training strategy and an implementation.
Last we select the parameters required by the algorithm.

\subfile{rq2/1_clustering_mutants}
\end{document}